---
title: "School failures on absence and study time"
author: "Samantha Gouveia"
date: "12/22/2023"
output:
 html_document:
 toc: yes
 toc_depth: 4
 toc_float: yes
 fig_width: 6
 fig_caption: yes
 number_sections: yes
 theme: readable
editor_options:
 chunk_output_type: console
---

```{=html}

<style type="text/css">

/* Cascading Style Sheets (CSS) is a stylesheet language used to describe the presentation of a document written in HTML or XML. it is a simple mechanism for adding style (e.g., fonts, colors, spacing) to Web documents. */

h1.title {  /* Title - font specifications of the report title */
  font-size: 24px;
  color: DarkRed;
  text-align: center;
  font-family: "Gill Sans", sans-serif;
}
h4.author { /* Header 4 - font specifications for authors  */
  font-size: 20px;
  font-family: system-ui;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - font specifications for the date  */
  font-size: 18px;
  font-family: system-ui;
  color: DarkBlue;
text-align: center;
}
h1 { /* Header 1 - font specifications for level 1 section title  */
    font-size: 22px;
    font-family: system-ui;
    color: navy;
    text-align: left;
}
h2 { /* Header 2 - font specifications for level 2 section title */
    font-size: 20px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - font specifications of level 3 section title  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - font specifications of level 4 section title  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}

body { background-color:white; }

.highlightme { background-color:yellow; }

p { background-color:white; }

</style>
```



# Introduction 
The data set used focuses on the alcohol consumption of student in secondary school. The data was collected through a survey of students in a math language course at two different schools. We are looking specifically at the correlation between study time and absences. 



# Data set 
 The data set has 32 independent variables and 1 dependent variable, the dependent variable being failure. The data will be focused on school Absences using simple linear regression.

```{r setup, include=FALSE}
s <- read.csv("https://raw.githubusercontent.com/sgouveia23/Assignment-2-data/main/student-mat.csv", header = TRUE )

   library(knitr)
   library(leaflet)
   library(EnvStats)
   library(MASS)
   library(phytools)
   library(boot)
   library(psych)
   library(tinytex)

   
   # Specifications of outputs of code in code chunks
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,  
    results = TRUE, comment = FALSE, options(digits = 2)   
                      )   
```

# Results

# Pair wise plot 

A pair wise scatter plot is being preformed to show the correlation between variables that do not already correlate with other independent variables.
```{r Pairwise scatterplot, echo = FALSE}


pairs.panels(s[,c(3,8,6)], pch=21, 
main="Pair-wise Scatter Plot of r numerical variables") #to show color grouping
# change c numbers 
```

# Check assumptions

 
```{r Assumptions, echo = FALSE}

ST <- s$studytime
SA <- s$absences

# Call lm() on price vs income
parametric_model <- lm(ST ~ SA)


#Plots to check for conditions
par(mfrow = c(1,2))
plot(parametric_model)
```
 When running a parametric test we must check for normality, constant variance. linearity, and influential points. The residual v fit plot shows a pattern, has all points in each section in a straight line, and shows that the data is not randomly distributed and do not have a constant variance. The normal Q-Q plot has a very noticeable right skew showing the data is not normally distributed. The scale-location plot shows the data is not spread out on an even spread. We see a influential point at 207.There is an outlier in all of the plots at 227 and 185.

# Bootstrap 

Bootstrap linear regression method was run over and over to find the distribution and the 95% confidence interval of the different regression models. 

```{r Bootstrap Regression Distribution, echo = FALSE}
  
beta0 <- numeric(1000)
beta1 <- numeric(1000)

#Create vectors same count as income observations
bootV <- 1:length(ST) 

#1000 bootstrap {replications
for(i in 1:1000){
  #Sample N for a bootstrap sample
  
  curS <- sample(bootV, length(ST), replace = TRUE)
  curA <- ST[curS]
  curT <- SA[curS]
  
  #Perform regression on that sample
  boot_reg <- lm(curA ~ curS)
  
  beta0[i] <- boot_reg$coefficients[1] 
  
  beta1[i] <- boot_reg$coefficients[2]  
}


#Plot Beta0 and Beta1 distributions
par(mfrow = c(1,2))
hist(beta0, main = "Bootstrap Sampled Intercepts", xlab = "Bootstrap Estimated Beta0s")
hist(beta1, main = "Bootstrap Sampled Slopes", xlab = "Bootstrap Estimated Beta1s")
```

When looking at the histogram we see the data is normally distributed.

# Confidence Interval

```{r Bootstrap Confidence Intervals, echo = FALSE}

  beta0_ci <- quantile(beta0, c(0.025, 0.975))
  beta1_ci <- quantile(beta1, c(0.025, 0.975))

```
We see that there is a correlation between study time and absences. By using the bootstrap method we were able to come to the conclusion of correlation using a non parametric way.

# Conclusion
 
Based on the results of the simple linear regressions tests run we see the bootstrap method is much better as it repeats the test over and over and therefore gives us a more accurate answer. We see this shown in the histograms as they show the data is normally distributed.

# References
Zubair, A(2023) "student-mat.csv "      https://www.kaggle.com/uciml/student-alcohol-consumption
